---
title: "Demo"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Demo}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE, eval = FALSE, 
  comment = "#>"
)
```

This `FILICA` R package conducts the Full Information Linked ICA algorithm (FI-LICA) to handle the missing data problem in multimodal fusion under the LICA framework.

# About 

In this file, we will demonstrate the implementation of FI-LICA, as well as LICA, in R, under both MAR and MCAR settings. 

The same version of this file with *printed-out results* can also bee seen at https://github.com/ruiyangli1/FILICA, named as `Demo.html`


*Note:*

* LICA is implemented using [MATLAB functions](https://fsl.fmrib.ox.ac.uk/fsl/fslwiki/FLICA), which will be called from R
* All the MATLAB functions and code scripts are in the folder `./MATLAB_code/flica` from our [GitHub website](https://github.com/ruiyangli1/FILICA)
  * MATLAB scripts to be called from R to run LICA are named as `code_*.m` (i.e., `code_completer.m`, `code_filica_update.m`, `code_nomiss.m`, `code_replace0.m`). ***Please make sure to change the path to `your path/FILICA/MATLAB_code/flica` in each script based on your folder location! Details in the User guideline section below.***
* ***Caution***: DO NOT run FI-LICA and completer case analysis at the same time as code will interfere with each other! 



# User guideline

* Create a new folder in your computer, named as `FILICA`, and set the working directory of R at the current `FILICA` folder
* Download the `./MATLAB_code/flica` folder from our [GitHub website](https://github.com/ruiyangli1/FILICA), and place it under the `FILICA` folder
* Open the following MATLAB scripts in the `FILICA/MATLAB_code/flica` folder: `code_completer.m`, `code_filica_update.m`, `code_nomiss.m`, `code_replace0.m`, and change the path in the first command line to **YOUR PATH`/FILICA/MATLAB_code/flica`** based on your folder location


# Examples

The following demonstrates how to use the functions in our FILICA package. 

## Load packages

```{r setup, message=FALSE,warning=FALSE}
sessionInfo()[[1]]$version.string

library(FILICA)
library(tidyverse)

library(scales)      # plot dF curve
breaks_log10 <- function(x) {
  low <- floor(log10(min(x)))
  high <- ceiling(log10(max(x)))
  10^(seq.int(low, high))
}
```


## MAR setting 

* n = 100
* miss 5% of data

### LICA on no-missing data
```{r fig.width = 7, fig.height = 5}
# -- generate data -- 
data = data_generate_MAR_ctnsH(nsubj = 100, seed = 7452, n_miss = 5) #--- n_miss doesn't matter b/c will only use no-miss data from data_generate()
# prepare data in lists
data = list(data$Y1, data$Y2)

# -- run LICA -- 
re_nomiss = LICA(data = data, ncomp = 5, niter = 1500, method = "nomiss")


# -- results -- 
## last dF
last_dF = re_nomiss$F.history[,length(re_nomiss$F.history)]-re_nomiss$F.history[,length(re_nomiss$F.history)-1]
message("Last dF = ", last_dF)
    
## H
H_nomiss = re_nomiss$H
H_nomiss_trans = t(H_nomiss); colnames(H_nomiss_trans) = paste0("est_comp", 1:ncol(H_nomiss_trans))
    
## concatenated XW
XWstack_nomiss = do.call(rbind, lapply(1:2, function(k){ 
  re_nomiss[["X"]][[k]][[1]] %*% diag( as.vector(re_nomiss[["W"]][[k]][[1]]) ) 
  }))
colnames(XWstack_nomiss) = paste0("est_comp", 1:5)

## dF curve
F_hist = t(re_nomiss$F.history)
F_hist_lag = lag(F_hist)
dF = data.frame(dF = F_hist - F_hist_lag)
dF$index = 1:nrow(dF)
dF %>% 
  ggplot(aes(x = index, y = dF)) + 
  geom_point(alpha = 0.5) + 
  geom_path() + 
  annotation_logticks() +
  scale_x_log10(breaks = breaks_log10,labels = trans_format(log10, math_format(10^.x))) +
  scale_y_log10(breaks = breaks_log10,labels = trans_format(log10, math_format(10^.x))) +
  labs(title = "dF curve") + 
  geom_hline(yintercept = 0.1, linetype = "dashed", color = "blue") + 
  theme_classic()
```

### LICA on data after replacing missing with 0
```{r fig.width = 7, fig.height = 5}
# -- generate data -- 
data = data_generate_MAR_ctnsH(nsubj = 100, seed = 7452, n_miss = 5) 
# prepare data in lists
data = list(data$Y1_miss, data$Y2_miss)

# -- run LICA -- 
re_replace0 = LICA(data = data, ncomp = 5, niter = 1500, method = "replace0")

# in case LICA stop early (e.g., n_miss=20), reduce flica_niter (by 25%) and rerun LICA until it works
# stop if reduce 18 times but still not work
flica_niter_used = 1500
stop_count = 1
while (class(re_replace0) == "try-error" & stop_count <= 18) {
  flica_niter_used = ceiling(0.75 * flica_niter_used)
  message("< Re-running LICA > ( ", flica_niter_used, " iterations )")
  re_replace0 = LICA(data = data, ncomp = 5, niter = flica_niter_used, method = "replace0")
  stop_count = stop_count + 1
}


# -- results -- 
## last dF
last_dF = re_replace0$F.history[,length(re_replace0$F.history)]-re_replace0$F.history[,length(re_replace0$F.history)-1]
message("Last dF = ", last_dF)
    
## H
H_replace0 = re_replace0$H
H_replace0_trans = t(H_replace0); colnames(H_replace0_trans) = paste0("est_comp", 1:ncol(H_replace0_trans))

## concatenated XW
XWstack_replace0 = do.call(rbind, lapply(1:2, function(k){ 
  re_replace0[["X"]][[k]][[1]] %*% diag( as.vector(re_replace0[["W"]][[k]][[1]]) ) 
  }))
colnames(XWstack_replace0) = paste0("est_comp", 1:5)

## dF curve
F_hist = t(re_replace0$F.history)
F_hist_lag = lag(F_hist)
dF = data.frame(dF = F_hist - F_hist_lag)
dF$index = 1:nrow(dF)
dF %>% 
  ggplot(aes(x = index, y = dF)) + 
  geom_point(alpha = 0.5) + 
  geom_path() + 
  annotation_logticks() +
  scale_x_log10(breaks = breaks_log10,labels = trans_format(log10, math_format(10^.x))) +
  scale_y_log10(breaks = breaks_log10,labels = trans_format(log10, math_format(10^.x))) +
  labs(title = "dF curve") + 
  geom_hline(yintercept = 0.1, linetype = "dashed", color = "blue") + 
  theme_classic()
```

### LICA on completers
```{r fig.width = 7, fig.height = 5}
# -- generate data -- 
data = data_generate_MAR_ctnsH(nsubj = 100, seed = 7452, n_miss = 5) 
# prepare data in lists
data = list(data$Y1_miss, data$Y2_miss)

# -- run LICA -- 
re_completer = LICA(data = data, ncomp = 5, niter = 1500, method = "completer")


# -- results -- 
## last dF
last_dF = re_completer$F.history[,length(re_completer$F.history)]-re_completer$F.history[,length(re_completer$F.history)-1]
message("Last dF = ", last_dF)

## H
H_completer = re_completer$H
H_completer_trans = t(H_completer); colnames(H_completer_trans) = paste0("est_comp", 1:ncol(H_completer_trans))

## concatenated XW
XWstack_completers = do.call(rbind, lapply(1:2, function(k){ 
  re_completer[["X"]][[k]][[1]] %*% diag( as.vector(re_completer[["W"]][[k]][[1]]) ) 
  }))
colnames(XWstack_completers) = paste0("est_comp", 1:5)

## dF curve
F_hist = t(re_completer$F.history)
F_hist_lag = lag(F_hist)
dF = data.frame(dF = F_hist - F_hist_lag)
dF$index = 1:nrow(dF)
dF %>% 
  ggplot(aes(x = index, y = dF)) + 
  geom_point(alpha = 0.5) + 
  geom_path() + 
  annotation_logticks() +
  scale_x_log10(breaks = breaks_log10,labels = trans_format(log10, math_format(10^.x))) +
  scale_y_log10(breaks = breaks_log10,labels = trans_format(log10, math_format(10^.x))) +
  labs(title = "dF curve") + 
  geom_hline(yintercept = 0.1, linetype = "dashed", color = "blue") + 
  theme_classic()
```

### FI-LICA

* Results from the completer cases analysis (`completer_results_saved_from_FILICA`) will also be stored at the Environment during the process. 
```{r fig.width = 7, fig.height = 3}
# -- generate data -- 
data = data_generate_MAR_ctnsH(nsubj = 100, seed = 7452, n_miss = 5) 
# prepare data in lists
data = list(data$Y1_miss, data$Y2_miss)

# -- run FI-LICA -- 
re_filica = FI_LICA(data = data, ncomp = 5, flica_niter = 1500, n = 20, flica_niter2 = 1000, rescale = TRUE, re_completer = NULL)
## or if want to use the saved results from completer case analysis: re_filica = FI_LICA(data = data, ncomp = 5, flica_niter = 1500, n = 20, flica_niter2 = 1000, rescale = TRUE, re_completer = re_completer)

# -- results --
## number of modality
mod_n = length(re_filica$results[["X"]])
## H
H_filica = re_filica$results$H
H_filica_trans = t(H_filica); colnames(H_filica_trans) = paste0("est_comp", 1:ncol(H_filica_trans))
## concatenated XW
XW_filica = do.call(rbind, lapply(1:mod_n, function(k){ 
  re_filica$results[["X"]][[k]][[1]] %*% diag( as.vector(re_filica$results[["W"]][[k]][[1]]) ) 
  }))
colnames(XW_filica) = paste0("est_comp", 1:5)

## convergence
par(mfrow=c(1,2))
plot(re_filica$H_convergence, ylab = "H_convergence", type="l")
plot(re_filica$XW_convergence, ylab = "XW_convergence", type="l")
```
```{r fig.width = 11, fig.height = 9}
## dF curve
F_hist = re_filica$F_hist
F_hist_lag = lag(F_hist)
dF = data.frame(F_hist - F_hist_lag)
colnames(dF) = paste0("update", 1:20)
dF$index = 1:nrow(dF)
dF %>% 
  pivot_longer(1:20, names_to = "update", values_to = "dF") %>% 
  mutate(update = factor(update,levels = paste0("update",1:20))) %>% 
  filter(update %in% c(paste0("update",1:20))) %>% 
  ggplot(aes(x = index, y = dF)) + 
  geom_point(alpha = 0.5) + 
  geom_path() + 
  facet_wrap(~update) +
  annotation_logticks() +
  scale_x_log10(breaks = breaks_log10,labels = trans_format(log10, math_format(10^.x))) +
  scale_y_log10(breaks = breaks_log10,labels = trans_format(log10, math_format(10^.x))) +
  labs(title = "dF curve in FI-LICA Step 2") + 
  geom_hline(yintercept = 0.1, linetype = "dashed", color = "blue") + 
  theme_classic()
```


## MCAR setting

* n = 100
* miss 5% of data
* same code as above, expect using `data_generate_MCAR`() instead of `data_generate_MAR_ctnsH()` for data generation
